Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions

### 2022-03-21 analysis/paper/shceof.Rmd
Agrega DMI

@@

@@

The connections between cEOFs and Tropical sources are also assessed.
Figure \@ref(fig:sst-psi-2) shows the regression maps of Sea Surface Temperatures (SST) and streamfunction anomalies at 200 hPa respectively upon standardised cEOF2, which can be interpreted as the anomalies associated with a 1 standard deviation increase of each index.
Besides showing the regression of the real and imaginary parts, we also show regressions in two intermediate directions. 
The Imaginary cEOF2 is associated with strong positive SST anomalies on the Central Pacific and negative anomalies over an area across the North of Australia and New Zealand, the South Pacific Convergence Zone (SPCZ) (Figure \@ref(fig:sst-psi-2).b1).
This pattern is almost canonically positive ENSO [@bamston1997] and indeed, the correlation between the Imaginary cEOF2 and the Oceanic Niño Index [@bamston1997] is significant and very high `r enso_cor[cEOF == "cEOF2" & part == "Imaginary"]$text`. 
On top of this ENSO-like pattern, there are positive anomalies in the western Indian Ocean and negative values in the eastern Indian Ocean, resembling a positive Indian Ocean Dipole.

>>>>

dmi <- download_dmi() %>%
  as.data.table() %>%
  .[, .(time = lubridate::as_datetime(Date), dmi = DMI)] %>%
  na.omit() %>%
  .[season(time) == "SON"] %>%
  .[, dmi := dmi*season_weights(time)] %>% 
  .[, .(dmi = mean(dmi)), by = .(time = seasonally(time))]

@@

```{r dmi_cor}
dmi_cor <-  ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm() %>%
  .[, hgt := scale(hgt), by = .(cEOF, part)] %>%
  dmi[., on = "time"] %>%
  na.omit() %>%
  .[, correlate(hgt, dmi), by = .(cEOF, part)]
```

@@

The connections between cEOFs and Tropical sources are also assessed.
Figure \@ref(fig:sst-psi-2) shows the regression maps of Sea Surface Temperatures (SST) and streamfunction anomalies at 200 hPa respectively upon standardised cEOF2, which can be interpreted as the anomalies associated with a 1 standard deviation increase of each index.
Besides showing the regression of the real and imaginary parts, we also show regressions in two intermediate directions. 
The Imaginary cEOF2 is associated with strong positive SST anomalies on the Central Pacific and negative anomalies over an area across the North of Australia and New Zealand, the South Pacific Convergence Zone (SPCZ) (Figure \@ref(fig:sst-psi-2).b1).
This pattern is almost canonically positive ENSO [@bamston1997] and indeed, the correlation between the Imaginary cEOF2 and the Oceanic Niño Index [@bamston1997] is significant and very high `r enso_cor[cEOF == "cEOF2" & part == "Imaginary"]$text`. 
On top of this ENSO-like pattern, there are positive anomalies in the western Indian Ocean and negative values in the eastern Indian Ocean, resembling a positive Indian Ocean Dipole. 
The correlation between the Imaginary cEOF2 and the DMI is `r dmi_cor[cEOF == "cEOF2" & part == "Imaginary"]$text`.
